6.867 Problem Set 3

### 2015-11-11 HW3/tex/HW3.tex

We are now interested in testing out neural network code on a real dataset. In our case, we're using a subset the very popular MNIST data collected by Yann Lecunn. This dataset is comprised of 28 by 28 pixel images of handwritten digits with corresponding labels that represent what numbers these images actually represent. For our features, each 28x28 images into a feature vector of 784 where each of these features represent the darkness of a pixel (0 means that its a white pixel, 1 means its a black pixel, everything in between are shades of grey).

@@

>>>>

We are now interested in testing out neural network code on a real dataset. In our case, we're using a subset the very popular MNIST data collected by Yann Lecunn. This dataset is comprised of 28 by 28 pixel images of handwritten digits with corresponding labels that represent what numbers these images actually represent. For our features, each 28x28 images into a feature vector of 784 where each of these features represent the darkness of a pixel (0 means that its a white pixel, 1 means its a black pixel, everything in between are shades of grey). Our dataset only consists of 6 distinct digits.

We again train both full-batch and stochastic gradient descent (batch size = 25) neural networks on the MNIST data, with a variety of different learning rates, regularization parameters and numbers of hidden nodes. The classification errors on both training and validation sets are found in Table 2, with best results in bold.

As expected, we see that SGD performs slightly worse than full-batch gradient descent, since it does not overfit as much. Our data seems to show that increasing the number of hidden nodes strictly increasing performance - this is not surprising, although we imagine that as you continue to increase $n_1$, you begin to get diminishing returns for increasing computational complexity. Decreasing the step size (i.e., smaller learning rate) also seems to lead us to achieve slightly better learning rates, although this causes the algorithm to require many more iterations to converge. A chart of the neural network classification error on the MNIST validation set can be found in Figure 2.

@@

\begin{figure}
\center
\includegraphics[height=4in, width=5in]{nnperf.jpg}
\caption*{Figure 2: Neural Network Classification Error on MNIST Validation Sets}
\end{figure}
